Observation of a Magnetic-Torque Enhanced by Interacting Spin Ensembles

### 2021-02-15 Mes papiers/Méca CR/18112020/SI/Meca_cross_SI.tex
tutu

\vspace{0.2in}
{\Large \hspace{1.6in}\textsc{Supplementary Material} }\\
\
\title{Resonant Dipole-Dipole Induced Mechanical Rotation of a Diamond Crystal}

@@

\section{NV$^-$ center Theory}
\subsection{NV spin hamiltonian}
The hamiltonian of the electronic spin of the negatively charged nitrogen-vacancy center in its ground state can be written as :
\begin{equation*}
  \hat{\mathcal{H}}_s=\hbar D S_z^2 + \hbar \gamma_e \textbf{B}\cdot\hat{\textbf{S}},
  \end{equation*} 
where $D = (2\pi) 2.87$ GHz is the crystal field splitting originating from spin-spin interactions, and $\gamma_e = 28 $GHz/T is the electron gyromagnetic ratio. We neglect contributions from the strain and local electric field in the hamiltonian since we are working with magnetic fields on the order of 100G. We also neglect the hyperfine interaction with the nuclear spin of the $^{14}$N atom since we are working with ensembles with a typical inhomogeneous broadening of 5 MHz. The \textbf{z} axis of the $S_z$ operator here is the axis formed by the nitrogen atom and the vacancy in the body frame.

@@

\subsection{Diamond crystalline axes and degeneracy conditions}
\begin{figure}[!ht]
  \centering \scalebox{0.45}{\includegraphics{SI_Cristallo}}
  \caption{a) Representation of the four crystalline axes of the diamond b) Representation of the magnetic field in the diamond crystalline basis c) Representation of the crystalline planes in the $(\theta , \phi)$ basis. The [110]$_\perp$ family of planes are shown using dashed lines, [100]$_\perp$ using plain lines. The [100] direction is marked by red circles, the [111] direction by green squares. The path of the magnetic field in the experiment of Fig. 3 of the main text is shown with blue dots.}
	\label{cristallo}
\end{figure}
There are four possible crystalline axis for the NV centers (so-called ``classes" of NV) which are depicted in Fig. \ref{cristallo} b) and correspond to the crystalline directions [$111$], [$1\bar 1 \bar 1$], [$\bar 1 1 \bar 1$] and [$\bar 1 \bar 1 1$].

The magnetic field is represented in the diamond basis in Fig. \ref{cristallo} a) where the usual polar and azimuthal angles $\theta$ and $\phi$ are defined with respect to the \textbf{z} ([$001$]) direction.

For some orientations of the magnetic field, the projection of the magnetic field on two or more NV axes will be identical, and therefore the energy level of the corresponding classes will be the same. These degeneracies are represented in Fig \ref{cristallo} c), where the dashed lines are the locus of the crystalline planes orthogonal to the [110] directions (6 planes in total). When the magnetic field belongs to these planes, we observe a degeneracy between two classes of NVs, as can be seen in the Fig. 4 of the main paper.

The plain lines are the loci of the planes orthogonal to the [100] directions (3 planes total). When the magnetic field angle lies in these planes, every class is resonant with another class, as can be seen in Fig. \ref{CR_22}.
The red circles correspond to the [100] directions, for which the four classes of NV are degenerate. The green squares correspond to the [111] direction where one class is aligned with the magnetic field, and the three others are degenerate. Finally the blue dots correspond to the path followed by the magnetic field in the experiment presented in Fig 3 of the main text, where we can see that a plane from the [110]$_\perp$ family is being crossed.

@@

\subsection{$T_1$ Measurement}
\begin{figure}[!ht]
  \centering \scalebox{1.2}{\includegraphics{T1_soustraction_shema}}
  \caption{$T_1$ measurement protocol. Green bars represent laser excitation, red bar represent resonant microwave $\pi$ pulse.}
	\label{T1_protocol}
\end{figure}

As shown in the Fig. 2-c) of the main text, the spin lifetime of the NV centers is modified in the presence of cross-relaxation with other classes of NV centers.
Here we present the protocol for removing the effects of charge state transfer in the dark, which masks the PL signal decay induced solely by spin depolarization. 
The protocol described in Fig. \ref{T1_protocol} consists in using two sequences. In the first sequence the spins are initially polarized in their $m_s=\ket{0}$ state through a 1 ms green laser excitation pulse and then left to evolve in the dark for a variable dark time $\tau$. The spin state is finally read out using a 10 $\mu$s laser pulse (shorter than the polarization time of the spins).

The second sequence uses the same parameters (polarization time, dark time and readout time) than the first sequence, but uses an extra resonant microwave $\pi$ pulse tuned to a transition of one of the four classes of NV$^-$ right before the readout pulse. The latter sequence brings the population of the $m_s=\ket{0}$ state to the $m_s=\ket{\pm 1}$ state of one class of NV centers.

Writing $S_1$ the signal obtained by the first sequence and $S_2$ the one obtained by the second sequence, and then measuring the difference $S_1-S_2$ between the two, we can measure the spin state population of a single class, and at the same time remove unwanted contributions to the photoluminescence, such as charge state transfer in the dark (which would be common to the two signals and therefore subtracted).
In order to avoid low frequency noises or drift, we alternate both sequence when performing the measurement.

@@

A permanent magnet is placed a few cm away from the diamond sample in order to apply a uniform magnetic field to the NV centers together with an electro-magnet.

To calibrate the magnetic field magnitude $B$, and its orientation $\theta$ with respect to the NV axis, we record Optically Detected Magnetic Resonance (ODMR) spectra and use two orientations to determine both the angle and the magnetic field amplitude.

@@

\subsection{Signal drift for levitating particles}

@@

The magnetic torque responsible for the motion of the diamond fundamentally comes from the crystalline anisotropy of the NV centers and from the transverse field $B_\perp$ responsible for mixing the eigenstates in the stationary state. We will start by considering the torque from a single NV center.
%The hamiltonian of a single NV under a magnetic field is 
%\be
%\hat{H}=\hbar D S_z^2+\hbar \gamma_e\bm S \cdot \bm B
%\ee
We simplify the problem by assuming that the motion is in the $x-z$ plane so the magnetic part of the hamiltonian reads 
$H_{B}= \hbar \gamma_e B  (\hat{S_x} \sin\theta + \hat{S_z} \cos\theta)$
where $\theta$ is the angle between the B field and NV center quantization axis.
We thus obtain the spin torque 
\be
\hat \tau_s = -\partial \hat{H} /\partial \theta = \hbar \gamma_e B (-\cos\theta \hat{S_x}+\sin\theta \hat{S_z}).
\ee
The mean value of the torque in terms of the reduced density matrix elements $\rho_{ij}$ in the basis of the $S_z$ eigenstates ($-1,0,1$) in the body fixed frame is
\be
\langle \hat \tau_s \rangle = \hbar \gamma_e B   (\rho_{-1-1}-\rho_{11})\sin\theta - \hbar \frac{\gamma_e B}{\sqrt{2}} S \cos\theta\label{torquemean}
\ee
where we introduced $S=\rho_{0,1}+\rho_{1,0}+\rho_{0,-1}+\rho_{-1,0}$.
The reduced density matrix elements are traced over a bath consisting of laser photons used to polarized the NV at a rate $\gamma_{\rm las}$, phonons or spin-fluctuators acting on the spin populations at a rate $R_1=1/T_1$ and 
$P_1$ or nuclear spins dephazing the electronic spin at a rate $1/T_2^*$.
In the limit $\gamma_e B \ll D$, and $\gamma_{\rm las}\gg \Gamma_1$ the laser still polarizes the spins in the ground state so that $\rho_{00}\gg {\rho_{11},\rho_{-1-1}}$.
Assuming that the pure dephazing $T_2^*$ is much shorter than the sum of the population relaxation $T_1/2$ and the laser induced repolarization $\gamma_{\rm las}$ the equations of motion for the coherences are

\bea
\frac{\partial \rho_{01}}{\partial t}&=&-\frac{1}{2T_2^*} \rho_{01}-i\frac{ \gamma_e B}{\sqrt{2}}\sin\theta-i\rho_{01}D+\mathcal{O}(\frac{(\gamma_e B)^2}{D})\\
\frac{\partial \rho_{0-1}}{\partial t}&=&-\frac{1}{2T_2^*}  \rho_{0-1}-i\frac{ \gamma_e B}{\sqrt{2}}\sin\theta-i\rho_{0-1}D+\mathcal{O}(\frac{(\gamma_e B)^2}{D}).
\eea
%Taking the complex value of these equations give the two other eq coherence terms. 
%The coherences can have an impact on the angle since they are responsible for the mean torque applied to the particle. 
The motional dynamics is very slow compared to the zero-field and magnetic field splittings. The latter are also larger then the decoherence rate $1/T_2^*$ so we can adiabatically eliminate the coherences and find 
\bea
\rho_{01}=\rho_{10}\approx-\frac{\gamma_e B \sin\theta}{\sqrt{2} D} \quad {\rm and} \quad \rho_{0-1}=\rho_{-10}\approx-\frac{\gamma_e B \sin\theta}{\sqrt{2} D}\eea
Since 
\bea
\rho_{11}-\rho_{-1-1}=\mathcal{O}(\frac{\gamma_e B}{D}).
\eea 
Re-injecting these expressions in the expression for the mean torque, we get 
\bea
\langle \hat \tau_s \rangle = \frac{\hbar(\gamma_e B)^2}{D}\sin2\theta
 +  \mathcal{O}(\frac{(\gamma_e B)^3}{D^2}).\label{torque_anal}
\eea

@@

Indeed, supposing that $\gamma B \ll D$, so that $H_{B}$ can be treated as a perturbation to the spin-spin hamiltonian, 
the perturbed energy $\epsilon_0$ of $\ket{0}$ is 
$$ \epsilon_g=\sum_{m_s=\pm 1}\frac{ |\bra{0} H_B \ket{\pm 1}|^2}{-\epsilon^0_{\pm 1}}=-\hbar\frac{(\gamma_e B)^2 }{D} \sin^2\theta
$$

The formula $ -\partial \epsilon_g /\partial \theta$ then gives Eq. \ref{torque_anal}. It is the equation that is used in the core of the manuscript. 
In the approximate regime of the present study, the Feynman-Hellman theorem that relates the angular derivative of the mean energy to the torque in the pure state case, 
can in fact also be applied directly here.

\begin{figure}[ht]
\centering{ \scalebox{0.4} {\includegraphics{map_torque_1_classe}}}
  \caption{\textbf{a)} Skecth showing an NV center aligned to the \textbf{z} direction in the presence of an external magnetic field \textbf{B} at an angle $\theta$ with respect the NV axis. The resulting spin vector $\langle \mathbf{ \hat S} \rangle$ of the NV center is shown by the blue arrow.
   \textbf{b)} Longitudinal ($S_\parallel$ in dashed line) and transverse ($S_\perp$ in plain line) components of the average value of the spin operator, in $\hbar$ unit, as a function of $\theta$ and for a magnetic field amplitude $|\mathbf{B}|=100$ G.   
   \textbf{c)} Amplitude of the magnetic torque acting on a single spin as a function of $\theta$ for $|\mathbf{B}|=100$ G.
   \textbf{d)} Amplitude of the same magnetic torque in the crystalline basis with $\theta$ and $\phi$ being the polar and azimuthal angle with respect to the [100] direction.}
  		\label{Torque1classe}
\end{figure}

Another way to estimate the torque is depicted in Fig \ref{Torque1classe}. Under green excitation and in the presence of an external magnetic field, the spins will acquire a magnetization $\gamma_e \langle\hat{\mathbf S}\rangle$ which, under the low magnetic fields ($<$ 200 G) we are working at, will be oriented at an angle of 90$^\circ$ from the NV axis : $\langle \hat S_z \rangle \approx 0$ and $\langle \hat S_\perp \rangle \neq 0$. This magnetization vanishes when the magnetic field is aligned with the NV center since there is no longer a transverse field responsible for the mixing of the eigenstates.

The magnetization of the NV center is therefore not aligned with the magnetic field, except when the field is also at a 90$^\circ$ angle from the NV axis, which means that the magnetic torque $\mathbf \Gamma = \gamma_e \langle\hat{\mathbf S}\rangle \times \mathbf B $ will be non-zero everywhere except when the field is aligned with the center, or in the plane normal to the direction of the center. We can describe each NV center as a paramagnetic defect with the anisotropic magnetic susceptibility $\chi = \begin{pmatrix}-\chi_\perp & 0 & 0\\ 0 & -\chi_\perp & 0 \\ 0 & 0 & 0  \end{pmatrix}$ in the $(\mathbf{x},\mathbf{y},\mathbf{z})$ basis where $\mathbf{z}$ is the orientation of the NV center.

@@

By using the inertia moment formula of a sphere : $I=\frac{2}{5}m r^2$, we can then rewrite the angular displacement as $$ \delta\theta = \frac{\langle\Gamma_{\mathrm{1 spin}}\rangle n(NV^-)}{\frac{2}{5}m_Cr^2\omega_T^2} \approx 10^{-4} rad$$ where $n(NV^-) \approx 5 \cdot 10^{-6}$ (5 ppm) is the number of NV centers per atoms in the crystal, $m_C \approx 2 \cdot 10^{-26}$ kg is the average weight of a carbon atom (we assume that the bulk of the diamond weight comes from carbon atoms), $r= 7.5$ $\mu$m is the typical radius of our diamonds and $\omega_T = 6.3\cdot 10^3$ rad/s is the typical value of the trap angular frequency.

>>>>

\vspace{0.2in}
{\Large \hspace{1.6in}\textsc{Supplementary Material} }\\
\
\title{Observation of a Magnetic-Torque Enhanced by Interacting Spin Ensembles}

@@

\section{NV$^-$ center Theory}
\subsection{NV spin hamiltonian}
The hamiltonian of the electronic spin of the negatively charged nitrogen-vacancy center in its ground state can be written as :
\begin{equation*}
  \hat{\mathcal{H}}_s=\hbar D S_{z'}^2 + \hbar \gamma_e \textbf{B}\cdot\hat{\textbf{S}},
  \end{equation*} 
where $D = (2\pi) 2.87$ GHz is the crystal field splitting originating from spin-spin interactions, and $\gamma_e = 28 $GHz/T is the electron gyromagnetic ratio. 
The \textbf{z'} axis in the $S_{z'}$ operator here is the axis formed by the nitrogen atom and the vacancy in the body fixed frame.
We neglect contributions from the strain and local electric field in the hamiltonian since we are working with magnetic fields on the order of 0.1T, which induce splittings  larger than the splitting of the zero-field ESR line ($\approx 20$ MHz). We also neglect the hyperfine interaction with the nuclear spin of the $^{14}$N atom since we are working with ensembles with typical inhomogeneous broadening of 5~MHz.

@@

\subsection{Diamond crystalline axes and degeneracy conditions}
\begin{figure}[!ht]
  \centering \scalebox{0.45}{\includegraphics{SI_Cristallo}}
  \caption{a) Representation of the four crystalline axes of the diamond b) Representation of the magnetic field in the diamond crystalline basis c) Representation of the crystalline planes in the $(\theta , \phi)$ basis. The [110]$_\perp$ family of planes are shown using dashed lines, [100]$_\perp$ using plain lines. The [100] direction is marked by red circles, the [111] direction by green squares. The magnetic field path in the experiment of Fig. 3 of the main text is shown with blue dots.}
	\label{cristallo}
\end{figure}
There are four possible crystalline axes for the N-V direction (so-called ``classes" of NV) in the diamond. They are depicted in Fig. \ref{cristallo} b) and correspond to the crystalline directions [$111$], [$1\bar 1 \bar 1$], [$\bar 1 1 \bar 1$] and [$\bar 1 \bar 1 1$].

The magnetic field direction is represented in Fig. \ref{cristallo} a), where the usual polar and azimuthal angles $\theta$ and $\phi$ are defined with respect to the \textbf{z'} ([$001$]) direction. For some orientations of the magnetic field, the projection of the magnetic field on two or more NV axes will be identical, and therefore the energy level of the corresponding classes will be the same. These degeneracies are represented in Fig \ref{cristallo} c), where the dashed lines are the locii of the crystalline planes orthogonal to the [110] directions (6 planes in total). When the magnetic field belongs to these planes, we observe a degeneracy between two classes of NVs, as can be seen in the Fig. 4 of the main paper.

The plain lines are the locii of the planes orthogonal to the [100] directions (3 planes in total). When the magnetic field lies in these planes, all classes are co-resonant, as can be seen in Fig. \ref{CR_22}.
The red circles correspond to the [100] directions, for which the four classes of NV are degenerate. The green squares correspond to the [111] direction where one class is aligned with the magnetic field, and the three others are degenerate. Finally the blue dots correspond to the path followed by the magnetic field in the experiment presented in Fig 3 of the main text, where we can see that a plane from the [110]$_\perp$ family is being crossed.

@@

\subsection{$T_1$ measurement}
\begin{figure}[!ht]
  \centering \scalebox{1.2}{\includegraphics{T1_soustraction_shema}}
  \caption{$T_1$ measurement protocol. Green bars represent laser excitation, red bar represent resonant microwave $\pi$ pulse.}
	\label{T1_protocol}
\end{figure}

As shown in the Fig. 2-c) of the main text, the spin lifetime of the NV centers is modified in the presence of cross-relaxation with other classes of NV centers.
Here we present the protocol for removing the effects of charge state transfer in the dark, which mask the PL signal decay induced solely by spin depolarization. 
The protocol described in Fig. \ref{T1_protocol} consists in using two sequences. In the first sequence the spins are initially polarized in the $\ket{m_s=0}$ state through a 1 ms green laser excitation pulse and then left to evolve in the dark for a variable dark time $\tau$. The spin state is finally read out using a 10 $\mu$s laser pulse (shorter than the polarization time of the spins).

The second sequence uses the same parameters (polarization time, dark time and readout time) than the first sequence, but uses an extra resonant microwave $\pi$ pulse tuned to a transition of one of the four classes of NV$^-$ right before the readout pulse. The latter sequence brings population from the $\ket{m_s=0}$ state to the $\ket{m_s=\pm 1}$ state for one class of NV centers.

By measuring the difference between the two signals obtained in these two measurements, we can extract the evolution of the spin state population from a single NV class and, at the same time, remove unwanted contributions to the photoluminescence, such as charge state transfer in the dark (which give the same background contribution to the to the measurements).
In order to avoid low frequency noises such as laser drifts from the focal point or intensity fluctuations, we alternate both sequences while performing the measurement.

@@

A neodymium permanent magnet and an electro-magnet are placed a few centimeters away from the diamond sample in order to apply a uniform and controllable magnetic field to the NV centers.

To calibrate the magnetic field magnitude $B$, and its orientation $\theta$ with respect to the NV axis, we record Optically Detected Magnetic Resonance (ODMR) spectra and record two the frequency of two transitions $|0 \rangle \rightarrow |-1\rangle$ and $|0 \rangle \rightarrow |+1\rangle$ from the same class to determine both the angle of the B field with respect to this class and the magnetic field amplitude.

@@

\subsection{Angular signal drift for levitating particles}

@@

The magnetic torque responsible for the motion of the diamond fundamentally comes from the anisotropy of the NV centers and from the transverse field $B_\perp$ responsible for mixing the eigenstates in the stationary state. We will start by considering the torque from a single NV center.
%The hamiltonian of a single NV under a magnetic field is 
%\be
%\hat{H}=\hbar D S_z^2+\hbar \gamma_e\bm S \cdot \bm B
%\ee
Without lack of generality, we will assume that the B field points in the $z$ direction and take the motion to be in the $x-z$ plane (in the lab frame), see Fig.~\ref{axes}.

\begin{figure}[ht]
\centering{ \scalebox{0.4} {\includegraphics{axes.eps}}}
  \caption{Notations used to define the axes in the body fixed and laboratory frames of reference $\mathcal{R'}$ and $\mathcal{R}$ respectively.}
  		\label{axes}
\end{figure}

In the body fixed frame, the magnetic part of the hamiltonian reads 
$\hat {H}_{B}= \hbar \gamma_e B  (\hat{S}_{x'} \sin\theta + \hat{S}_{z'} \cos\theta)$
where $\theta$ is the angle between the B field and NV center quantization axis $z'$.
We thus obtain the spin torque operator
\be
\hat \tau_s = -\frac{\partial \hat{H}}{\partial \theta} = \hbar \gamma_e B (-\cos\theta \hat{S}_{x'}+\sin\theta \hat{S}_{z'}).
\ee
The mean value of the torque operator in terms of the reduced density matrix elements $\rho_{ij}$ in the basis of the $S_{z'}$ eigenstates $|-1\rangle_{z'},|0\rangle_{z'},|1\rangle_{z'}$ is 
\be
\langle \hat \tau_s \rangle =\mathrm{Tr}_\mathcal{B}(\hat \rho \hat \tau_s) = \hbar \gamma_e B   (\rho_{-1-1}-\rho_{11})\sin\theta - \hbar \frac{\gamma_e B}{\sqrt{2}} S \cos\theta\label{torquemean}
\ee
where we introduced $S=\rho_{0,1}+\rho_{1,0}+\rho_{0,-1}+\rho_{-1,0}$.
The bath $\mathcal{B}$ over which the trace is performed consists of laser photons used to polarized the NV at a rate $\gamma_{\rm las}$, phonons or spin-fluctuators acting on the spin populations at a rate $\Gamma_1=1/T_1$ and 
$P_1$ centers or nuclear spins dephazing the electronic spin at a rate $1/T_2^*$.
In the limit $\gamma_e B \ll D$, and $\gamma_{\rm las}\gg \Gamma_1$ the laser efficiently polarizes the electronic spins in the ground state so that $\rho_{00}\gg {\rho_{11},\rho_{-1-1}}$.
The pure dephazing $T_2^*\approx 100$ns is much shorter than the sum of the population relaxation time $T_1/2 \le 1$ms and the laser induced repolarization time $1/\gamma_{\rm las} \le $ 100 $\mu$s. The equations of motion for the coherences thus read

\bea
\frac{\partial \rho_{01}}{\partial t}&=&-\frac{1}{2T_2^*} \rho_{01}-i\frac{ \gamma_e B}{\sqrt{2}}\sin\theta-i\rho_{01}D+\mathcal{O}(\frac{(\gamma_e B)^2}{D})\\
\frac{\partial \rho_{0-1}}{\partial t}&=&-\frac{1}{2T_2^*}  \rho_{0-1}-i\frac{ \gamma_e B}{\sqrt{2}}\sin\theta-i\rho_{0-1}D+\mathcal{O}(\frac{(\gamma_e B)^2}{D}).
\eea
%Taking the complex value of these equations give the two other eq coherence terms. 
%The coherences can have an impact on the angle since they are responsible for the mean torque applied to the particle. 
The characteristic motional dynamics is very slow compared to the zero-field and magnetic field rates $D$ and $\gamma B$. The latter are also much larger then the decoherence rate $1/T_2^*$ in our experiments, so we can adiabatically eliminate the coherences and find 
\bea
\rho_{01}=\rho_{10}\approx-\frac{\gamma_e B \sin\theta}{\sqrt{2} D} \quad {\rm and} \quad \rho_{0-1}=\rho_{-10}\approx-\frac{\gamma_e B \sin\theta}{\sqrt{2} D}.\eea
Since 
\bea
\rho_{11}-\rho_{-1-1}=\mathcal{O}(\frac{\gamma_e B}{D}).
\eea 
Re-injecting these expressions in the expression for the mean torque, we get 
\bea
\langle \hat \tau_s \rangle = \frac{\hbar(\gamma_e B)^2}{D}\sin2\theta
 +  \mathcal{O}(\frac{(\gamma_e B)^3}{D^2}).\label{torque_anal}
\eea

@@

Indeed, supposing that $\gamma B \ll D$, so that $H_{B}$ can be treated as a perturbation to the spin-spin hamiltonian $D \hat S_{z'}^2$, 
the perturbed energy $\epsilon_0$ of $\ket{0}$ is 
\bea \epsilon_g=\sum_{m_s=\pm 1}\frac{ |\bra{0} H_B \ket{\pm 1}|^2}{-\epsilon^0_{\pm 1}}=-\hbar\frac{(\gamma_e B)^2 }{D} \sin^2\theta
\eea

Taking $ -\partial \epsilon_g /\partial \theta$ then gives Eq. \ref{torque_anal}. It is the equation that is used in the core of the manuscript. 
In the approximate regime of the present study, the Hellmann-Feynman theorem (exact for pure states) that relates the angular derivative of the mean energy to the torque is correct in the above-described limits where dissipation is negligible. \\

\begin{figure}[ht]
\centering{ \scalebox{0.4} {\includegraphics{map_torque_1_classe}}}
  \caption{\textbf{a)} Skecth showing an NV center aligned to the \textbf{z'} direction in the presence of an external magnetic field \textbf{B} at an angle $\theta$ with respect the NV axis. The resulting spin vector $\langle \mathbf{ \hat S} \rangle$ of the NV center is shown by the blue arrow.
   \textbf{b)} Longitudinal ($S_\parallel$ in dashed line) and transverse ($S_\perp$ in plain line) components of the average value of the spin operator, in $\hbar$ unit, as a function of $\theta$ and for a magnetic field amplitude $|\mathbf{B}|=100$ G.   
   \textbf{c)} Amplitude of the magnetic torque acting on a single spin as a function of $\theta$ for $|\mathbf{B}|=100$ G.
   \textbf{d)} Amplitude of the same magnetic torque in the crystalline basis with $\theta$ and $\phi$ being the polar and azimuthal angle with respect to the [100] direction.}
  		\label{Torque1classe}
\end{figure}

Another way to estimate the torque is depicted in Fig \ref{Torque1classe}. Under green excitation and in the presence of an external magnetic field, the spins will acquire a magnetization $\gamma_e \langle\hat{\mathbf S}\rangle$ which, under the low magnetic fields ($<$ 200 G) we are working at, will be oriented at an angle of 90$^\circ$ from the NV axis : $\langle \hat S_{z'} \rangle \approx 0$ and $\langle \hat S_\perp \rangle \neq 0$. This magnetization vanishes when the magnetic field is aligned with the NV center since there is no longer a transverse field responsible for the mixing of the eigenstates.

The magnetization of the NV center is therefore not aligned with the magnetic field, except when the field is also at a 90$^\circ$ angle from the NV axis, which means that the magnetic torque $\mathbf \Gamma = \gamma_e \langle\hat{\mathbf S}\rangle \times \mathbf B $ will be non-zero everywhere except when the field is aligned with the center, or in the plane normal to the direction of the center. We can describe each NV center as a paramagnetic defect with the anisotropic magnetic susceptibility $\chi = \begin{pmatrix}-\chi_\perp & 0 & 0\\ 0 & -\chi_\perp & 0 \\ 0 & 0 & 0  \end{pmatrix}$ in the $(\mathbf{x'},\mathbf{y'},\mathbf{z'})$ basis where $\mathbf{z'}$ is the orientation of the NV center.

@@

By using the inertia moment formula of a sphere : $I=\frac{2}{5}m r^2$, we can then rewrite the angular displacement as $$ \delta\theta = \frac{\langle\Gamma_{\mathrm{1 spin}}\rangle n(NV^-)}{\frac{2}{5}m_Cr^2\omega_T^2} \approx 10^{-3} rad$$ where $n(NV^-) \approx 5 \cdot 10^{-6}$ (5 ppm) is the number of NV centers per atoms in the crystal, $m_C \approx 2 \cdot 10^{-26}$ kg is the average weight of a carbon atom (we assume that the bulk of the diamond weight comes from carbon atoms), $r= 7.5$ $\mu$m is the typical radius of our diamonds and $\omega_T = 6.3\cdot 10^3$ rad/s is the typical value of the trap angular frequency.
